The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-29 paper/paper.Rnw
Finshing last mode

<<default_conf>>=
  library(matrixStats)
  library(tidyverse)
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
  library(countrycode) # For region variable
  library(car)
  library(SAScii)
  library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
  library(lme4)
  library(modelr)
  library(ggrepel)
  library(xtable)
  library(ggthemes)

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  left_join(school_data_ready) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = selage) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

@@

library(brms)
# This linear model explains the pooled 90/10 differences
mod_tracking <-
  brms::brm(
    difference ~
      ztrack +
      zvoc +
      ztrack:zvoc, family = gaussian(),
  data = ready_data_age,
  warmup = 1000, iter = 2000, chains = 5
  )

mod_tracking
bayes_R2(mod_tracking)
marginal_effects(mod_tracking)

@@

# Check whether the previous model can be generalized to a multilevel
# and see how much it can explain between country var
mod_tracking <-
  brms::brm(
    difference ~
      ztrack +
      zvoc +
      ztrack:zvoc + (1 | country),
  family = gaussian(),
  data = ready_data_age,
  prior = c(set_prior("normal(0, 1)", coef = "ztrack"),
            set_prior("normal(0, 1)", coef = "zvoc")),
  warmup = 1000, iter = 2000, chains = 4
  )

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model

# The code bloew needs to be adapted to the model from above (
# this is from previous models)

new_data <-
  ready_data_age %>%
  filter(country %in% countries) %>%
  expand(country, age_selection, num_tracks) %>%
  filter(complete.cases(.))

predictions <- posterior_predict(mod_tracking, newdata = new_data)
sd_obs <- colSds(predictions)

@@

new_data %>%
  left_join(sum_ready_data) %>%
  filter(country %in% countries) %>%
  ggplot(aes(num_tracks %>% as.factor, avg_diff, fill = num_tracks)) +
  geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
  geom_point(colour = "red") +
  scale_y_continuous(limits = c(0, 4)) +
  geom_col(aes(y = pred, alpha = 0.4)) +
  geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
  facet_wrap(country ~ age_selection) +
  scale_fill_wsj()

y_rep <- posterior_predict(mod_tracking)

replication <-
  map(1:2000, ~ {
  ready_data %>%
  mutate(pred_sample = y_rep[sample(1:4000, 1), , drop = TRUE]) %>%
  summarize(perc_sample = mean(pred_sample > difference),
            minus_sample = mean(pred_sample - difference))
})

replication %>%
  enframe() %>%
  unnest() %>%
  qplot(perc_sample, geom = "histogram", data = .)

replication %>%
  enframe() %>%
  unnest() %>%
  summarize(mean(perc_sample < 0.5))

pp_check(mod_tracking, type = "error_binned", nsamples = 1000)
@

@@

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ ztrack + zvoc,
    family = gaussian(),
    data = .,
    warmup = 1000, iter = 2000, chains = 5
  )

@@

ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  mutate(pred_median = posterior_predict(mod_tracking) %>% colMedians()) %>%
  select(country, cum_diff, pred_median) %>%
  gather(cat, value, -country) %>%
  ggplot(aes(country, value, fill = cat)) +
  geom_col(position = "dodge") +
  coord_flip()

>>>>

<<default_conf>>=
  library(matrixStats)
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
  library(countrycode) # For region variable
  library(car)
  library(SAScii)
  library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
  library(lme4)
  library(modelr)
  library(ggrepel)
  library(xtable)
  library(ggthemes)
  library(tidyverse)

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = selage) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

@@

library(brms)
# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences

mod_tracking <-
  brms::brm(
    difference ~
      ztrack +
      zvoc +
      ztrack:zvoc, family = gaussian(),
  data = ready_data_age,
  warmup = 1000, iter = 2000, chains = 5
  )

mod_tracking

@@

bayes_R2(mod_tracking)
marginal_effects(mod_tracking,
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

# I don't know how to estimate VIF's with bayesian reasoning
model_formula <- difference ~ num_tracks + age_selection + zvoc + I(length > 0)
mod_tracking_ext <-
  brms::brm(model_formula, family = gaussian(),
  data = ready_data_age,
  warmup = 1000, iter = 2000, chains = 5
  )

bayes_R2(mod_tracking_ext)

# This is the closest I could get to getting vif's:
lm(model_formula, data = ready_data_age) %>%
  vif()

new_data <-
  ready_data_age %>%
  with(crossing(country = countries,
                num_tracks = unique(num_tracks),
                age_selection = unique(age_selection),
                zvoc = mean(zvoc, na.rm = TRUE),
                length = typical(length > 0)))

predictions <- posterior_predict(mod_tracking_ext, newdata = new_data)
sd_obs <- colSds(predictions)

@@

new_data %>%
  left_join(sum_ready_data) %>%
  ggplot(aes(num_tracks %>% as.factor, avg_diff, fill = num_tracks)) +
  geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
  geom_point(colour = "red") +
  scale_y_continuous(limits = c(0, 4)) +
  geom_col(aes(y = pred, alpha = 0.4)) +
  geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
  facet_wrap(country ~ age_selection) +
  scale_fill_wsj()
@

@@

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ I(age_selection >= 15),
    family = gaussian(),
    data = .,
    warmup = 1000, iter = 2000, chains = 5
  )

@@

m_data <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  mutate(pred_median = posterior_predict(mod_tracking) %>% colMeans(),
         ratio_pred = pred_median/cum_diff) %>%
  select(country, ratio_pred, cum_diff, pred_median)

m_data %>%
  ggplot(aes(ratio_pred)) +
  geom_histogram(bins = 50)

m_data %>%
  filter(ratio_pred < 1) %>%
  ggplot(aes(reorder(country, -ratio_pred), 1 - ratio_pred)) +
  geom_col() +
  scale_y_continuous(breaks = seq(0, 0.30, 0.05), limits = c(0, 0.3)) +
  coord_flip()

# Greece, France would have a decrease of about 20-25% of their cumulative
# inequality when adjusting for the age of selection.

# There is both under and over prediction but the over predictions are very extreme
